---
title: "fuelEconomy"
author: "Steve Zhang"
date: "July 3, 2017"
output: html_document
---



## R Markdown

Analysis of United States fuel economy dataset courtesy of the U.S. Department of Energy and the U.S. Environmental Protection Agency.

```{r}
vehicles = read.csv("vehicles.csv", stringsAsFactors = FALSE)
```

```{r}
labels = do.call(rbind, strsplit(readLines("varlabels.txt"), " - "))
```
```{r}
names(vehicles)
```
```{r}
cat("First year in dataset:", min(vehicles$year))
cat("Most recent year in dataset:", max(vehicles$year))
cat("Number of unique years in dataset:", length(unique(vehicles$year)))
```

Create a table of fuel types used by the vehicles in the dataset.

```{r}
table(vehicles$fuelType1)
```

Clean up transmission data and set to either Automatic or Manual
```{r}
vehicles$trany[vehicles$trany == ""] = NA
vehicles$trany2 = ifelse(substr(vehicles$trany, 1, 4) == "Auto", "Auto", "Manual")
vehicles$trany2 = as.factor(vehicles$trany2)
table(vehicles$trany2)
```

Use ddply to group by year and compute MPG for highway, city and combined.

```{r}
library(plyr)
library(ggplot2)
library(reshape2)
```


```{r}
mpgYear = ddply(vehicles, ~year, summarise, avgMPG = mean(comb08), avgHwy = mean(highway08),
                avgCity = mean(city08))
```

```{r}
ggplot(mpgYear, aes(year, avgMPG)) + geom_point() + geom_smooth() +
    xlab("Year") + ylab("Average MPG") + ggtitle("All Cars")
```
```{r}
gasCars = subset(vehicles, fuelType1 %in% c("Regular Gasoline", "Premium Gasoline", "Midgrade Gasoline") & fuelType2 == "" & atvType != "Hybrid")
```

```{r}
gasCars$displ = as.numeric(gasCars$displ)
ggplot(gasCars, aes(displ, comb08)) + geom_point() + geom_smooth()
```
Make a graph of average displacement vs. year.

```{r}
avgCarSize = ddply(gasCars, ~year, summarise, avgDispl = mean(displ))
ggplot(avgCarSize, aes(year, avgDispl)) + geom_point() + geom_smooth() +
    xlab("Year") + ylab("Average Engine displacement (L)")
```

```{r}
byYear = ddply(gasCars, ~year, summarise, avgMPG = mean(comb08), avgDispl=mean(displ))
```

Convert byYear data frame from wide format to long format with melt function.
```{r}
byYear2 = melt(byYear, id = "year")
levels(byYear2$variable) = c("Average MPG", "Avg Engine Displacement")
```


```{r}
ggplot(byYear2, aes(year, value)) + geom_point() + geom_smooth() + 
    facet_wrap(~variable, ncol=1, scales="free_y") + xlab("Year") + ylab("")
```
```{r}
gasCars4 = subset(gasCars, cylinders == "4")
ggplot(gasCars4, aes(factor(year), comb08)) + geom_boxplot() + facet_wrap(~trany2, ncol = 1) + theme(axis.text.x = element_text(angle=45)) + labs(x="Year", y="MPG")
```
```{r}
ggplot(gasCars4, aes(factor(year), fill=factor(trany2))) + geom_bar(position = "fill") + labs(x="Year", y="Proportion", fill="Transmission") + theme(axis.text.x = element_text(angle = 45)) + geom_hline(yintercept = 0.5, linetype=2)
```
```{r}
carsMake = ddply(gasCars4, ~year, summarise, numberOfMakes = length(unique(make)))
ggplot(carsMake, aes(year, numberOfMakes)) + geom_point() + labs(x="Year", y="Number Makes") +ggtitle("Four cylinder cars")
```

Makes available during every year of the study.
```{r}
uniqMakes = dlply(gasCars4, ~year, function(x) unique(x$make))
commonMakes = Reduce(intersect, uniqMakes)
commonMakes
```

```{r}
carsCommonMakes4 = subset(gasCars4, make %in% commonMakes)
avgMPG_common = ddply(carsCommonMakes4, ~year + make, summarise, avgMPG=mean(comb08))
ggplot(avgMPG_common, aes(year, avgMPG)) + geom_line() + facet_wrap(~make, nrow = 3)
```

```{r}
commonCars = subset(vehicles, make %in% commonMakes)
avgMPG_common = ddply(commonCars, ~year + make, summarise, avgMPG=mean(comb08))
ggplot(avgMPG_common, aes(year, avgMPG)) + geom_line() + facet_wrap(~make, nrow = 3)
```


