---
title: "NFLSimulation"
author: "Steve Zhang"
date: "July 6, 2017"
output: html_document
---

```{r}
library(ggplot2)
library(XML)
library(RSQLite)
library(stringr)
```
```{r}
year = 2013
offense = read.csv("offense.csv", stringsAsFactors = FALSE)
head(offense)
```
```{r}
defense = read.csv("defense.csv", stringsAsFactors = FALSE)
head(defense)
```

Clean up offense data by changing to appropriate data types and convert time of possesion to seconds from minutes.
```{r}
offense[, 1] = as.character(offense[, 1])
offense[, 2:13] = apply(offense[, 2:13], 2, as.numeric)
offense[, 14] = as.numeric(substr(offense[, 14], 1, 2)) * 60 
```

Clean up defense data in similar way.
```{r}
defense[, 1] = as.character(defense[, 1])
defense[, 2:13] = apply(defense[, 2:13], 2, as.numeric)
```

Merge offense and defense data frames and clean up columns.
```{r}
combine = merge(offense, defense, by.x="Team", by.y = "Team")
```

```{r}
colnames(combine)[2] = "Games"
colnames(combine)[3] = "OffPPG"
colnames(combine)[4] = "OffYPG"
colnames(combine)[5] = "OffPassYPG"
colnames(combine)[6] = "OffRushYPG"
combine$G.y = NULL
colnames(combine)[15] = "DefPPG"
colnames(combine)[16] = "DefYPG"
colnames(combine)[17] = "DefRushYPG"
colnames(combine)[18] = "DefPassYPG"
```
```{r}
combine$X = NULL
```

```{r}
hist(combine$OffPPG, breaks = 10, main="Offensive PPG", xlab="PPG", ylab="Number Teams")
```
```{r}
summary(combine$OffPPG)
```

```{r}
hist(combine$DefPPG, breaks = 10, main="Defensive PPG", xlab = "PPG", ylab = "Number Teams")
```
```{r}
hist(combine$X1stD.G, breaks=10, main="Offensive 1st Downs/Game",
     xlab="First Downs", ylab="Number Teams")
```

Create bar chart for teams offensive PPG.
```{r}
ppg = transform(combine, Team=reorder(Team, combine$OffPPG))
ggplot(ppg, aes(x=Team, y=OffPPG)) + geom_bar(stat="identity", color="black", fill="blue")+
    coord_flip() + labs(x="Team", y="OffPPG") + ggtitle("Average Points Per Game") +
    theme(plot.title = element_text(size=18, face="bold"))
```

```{r}
ypg = transform(combine, Team=reorder(Team, -combine$DefYPG))
ggplot(ypg, aes(x=Team, y=DefYPG)) + geom_bar(stat="identity", color="black", fill="blue")+
    coord_flip() + labs(x="Team", y="Avg Yards PPG") + ggtitle("Avg Yards Allowed") +
    theme(plot.title = element_text(size=18, face="bold"))
```

```{r}
ggplot(combine, aes(x=OffYPG, y=OffPPG)) + geom_point(shape=5, size=2) + geom_smooth() +
    labs(x="Yards Per Game", y="Points Per Game") + ggtitle("Offense YPG vs. PPG") +
    theme(plot.title = element_text(size=18, face="bold"))
```

```{r}
ggplot(combine, aes(x=DefYPG, y=DefPPG)) + geom_point(size=2)+geom_smooth() +
    labs(x="Yards Allowed per game", y="Points Allowed Per game") + ggtitle("Defense Yards vs. Points per Game") + theme(plot.title = element_text(size=15, face="bold"))
```

```{r}
combine$TOP = combine$TOP + rnorm(dim(combine)[1], mean = 0.0, sd=20)
```

```{r}
ggplot(combine, aes(x=TOP, y=OffPPG)) + geom_point(size=2) + geom_smooth()+
    labs(x="Time of Possession", y="PPG") + ggtitle("Time of Possession vs. Points Per Game") + theme(plot.title = element_text(size=15, face="bold"))
```

Create indices for offensive strength.
```{r}
offense$OPassStrength = max(offense[, 5]) - offense[, 5]
offense$OPassStrength = (1 - (offense$OPassStrength/max(offense$OPassStrength)))*100
offense$ORushStrength = max(offense[, 6]) - offense[, 6]
offense$ORushStrength = (1 - (offense$ORushStrength/max(offense$ORushStrength)))*100
offense$OPPGStrength = max(offense[, 3]) - offense[, 3]
offense$OPPGStrength = (1 - (offense$OPPGStrength/max(offense$OPPGStrength)))*100
offense$OYPGStrength = max(offense[, 4]) - offense[, 4]
offense$OYPGStrength = (1 - (offense$OYPGStrength/max(offense$OYPGStrength)))*100
```

Create consolidated measure of offensive strength
```{r}
offense$OffStrength = (offense$OPassStrength + offense$ORushStrength + offense$OPPGStrength + offense$OYPGStrength)/4
```

Create measures of defensive strength
```{r}
defense$DPassStrength = max(defense[, 6]) - defense[, 6]
defense$DPassStrength = defense$DPassStrength/max(defense$DPassStrength)*100
defense$DRushStrength = max(defense[, 5]) - defense[, 5]
defense$DRushStrength = defense$DRushStrength/max(defense$DRushStrength)*100
defense$DPPGStrength = max(defense[, 3]) - defense[, 3]
defense$DPPGStrength = defense$DPPGStrength/max(defense$DPPGStrength)*100
defense$DYPGStrength = max(defense[, 4]) - defense[, 4]
defense$DYPGStrength = defense$DYPGStrength/max(defense$DYPGStrength)*100
```

```{r}
defense$DefStrength = (defense$DPassStrength + defense$DRushStrength + defense$DPPGStrength + defense$DYPGStrength)/4
```

```{r}
simGame = function(home_team, away_team){
    off_game = subset(offense, Team==home_team | Team == away_team)[, c(1, 15, 16, 19)]
    def_game = subset(defense, Team  == home_team | Team == away_team)[, c(1, 15, 16, 19)]
    game = merge(off_game, def_game, by.x="Team", by.y="Team")
    home_off_net = game[game$Team == home_team, ]$OffStrength - game[game$Team==away_team, ]$DefStrength
    home_points = rnorm(1, mean = home_off_net, sd = 50)
    away_off_net = game[game$Team == away_team, ]$OffStrength - game[game$Team== home_team, ]$DefStrength
    away_points = rnorm(1, mean = away_off_net, sd = 50)
    if (home_points > away_points) {
        return(list(winner=home_team, loser=away_team))   
    } else {
        return(list(winner=away_team, loser=home_team))
    }
}
```

```{r}
simSeason = function() {
    #games_per_team = 16
    #create a schedule of games for the specified number of games per team
   # for (week in 1:games_per_team) {
    #    home_index = sample(1:32, 16, replace = FALSE)
     #   home_teams = data.frame(HomeTeam = offense[home_index, 1])
      #  away_teams = data.frame(AwayTeam = offense[-home_index, 1])
       # if (week == 1) {
        #    schedule = cbind(Week=week, HomeTeam=home_teams, AwayTeam = away_teams)
        #} else {
        #    temp = cbind(Week=week, HomeTeam=home_teams, AwayTeam=away_teams)
         #   schedule = rbind(schedule, temp)
        #}
    #}
    schedule = read.csv("2013games.csv", stringsAsFactors = FALSE)
    names(schedule) = c("HomeTeam", "AwayTeam")
    records = data.frame(Team=offense$Team)
    records$Wins = 0
    records$Losses = 0
    for(i in 1:nrow(schedule)) {
        home_team = schedule[i, 1]
        away_team = schedule[i, 2]
        n = length(home_team)
        winners = rep(NA, n)
        losers = rep(NA, n)
        for (j in 1:n) {
            gameResult = simGame(home_team[j], away_team[j])
            winners[j] = gameResult$winner
            losers[j] = gameResult$loser
        }
        if (i == 1) {
            results = data.frame(winner=winners, loser=losers)
        } else {
            temp = data.frame(winner=winners, loser=losers)
            results = rbind(results, temp)
        }
    }
    n = nrow(results)
    for (i in 1:n) {
        row = results[i, ]
        winTeam = row$winner
        loseTeam = row$loser
        records[records$Team==winTeam, ]$Wins = records[records$Team == winTeam, ]$Wins + 1
        records[records$Team==loseTeam, ]$Losses = records[records$Team == loseTeam, ]$Losses + 1
    }
    return(records)
}
```

```{r}
sim2013 = simSeason()
```


