---
title: "stockSim"
author: "Steve Zhang"
date: "July 10, 2017"
output: html_document
---

```{r}
library(XML)
library(ggplot2)
library(plyr)
library(reshape2)
library(zoo)
```
```{r}
url_to_open <-
'http://finviz.com/export.ashx?v=152&c=0,1,2,3,4,5,6,7,8,9,10,
11,12,13,14,15,16,17,18,19,20,21,22,23,24,25,26,27,28,29,30,31
,32,33,34,35,36,37,38,39,40,41,42,43,44,45,46,47,48,49,50,51,5
2,53,54,55,56,57,58,59,60,61,62,63,64,65,66,67,68'
```

```{r}
finviz = read.csv("finviz.csv", stringsAsFactors = FALSE)
```

```{r}
clean_numeric = function(s) {
    s = gsub("%|\\$|,|\\)|\\(", "", s)
    s = as.numeric(s)
}
```

Remove punctuation from numeric fields with clean_numeric function.

```{r}
finviz = cbind(finviz[, 1:6], apply(finviz[, 7:68], 2, clean_numeric))
```

```{r}
hist(finviz$Price, breaks = 100, main="Price Distribution", xlab="Price")
```

Put cap on x-axis and make histogram.
```{r}
hist(finviz$Price[finviz$Price < 150], breaks = 100, main="Price Distribution", xlab = "Price")
```

```{r}
sectorAvgPrice = aggregate(Price~Sector, data = finviz, FUN = "mean")
colnames(sectorAvgPrice)[2] = "Sector_Avg_Price"
ggplot(sectorAvgPrice, aes(x=Sector, y=Sector_Avg_Price, fill=Sector)) + geom_bar(stat = "identity") + ggtitle("Sector Avg Prices") + theme(axis.text.x = element_text(angle = 90, hjust = 1))
```
```{r}
industryAvg = aggregate(Price~Sector+Industry, data = finviz, FUN = "mean")
industryAvg = industryAvg[order(industryAvg$Sector, industryAvg$Industry),]
colnames(industryAvg)[3] = "avgPrice"
industryChart = subset(industryAvg, Sector == "Financial")
```

```{r}
ggplot(industryChart, aes(x=Industry, y=avgPrice, fill=Industry)) + geom_bar(stat="identity") + theme(legend.position = "none") + ggtitle("Industry Avg Prices") + theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

Investigate outlier price in the Property & Casualty Insurance industry.

```{r}
compChart = subset(finviz, Industry == "Property & Casualty Insurance")
```

```{r}
ggplot(compChart, aes(x=Company, y=Price, fill=Company)) + geom_bar(stat="identity") + theme(legend.position = "none") + ggtitle("Company Avg Prices") + theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

Remove Berkshire Hathaway outlier from dataset.
```{r}
finviz = subset(finviz, Ticker != "BRK-A")
sectorAvgPrice = aggregate(Price~Sector, data = finviz, FUN = "mean")
colnames(sectorAvgPrice)[2] = "Sector_Avg_Price"
ggplot(sectorAvgPrice, aes(x=Sector, y=Sector_Avg_Price, fill=Sector)) + geom_bar(stat = "identity") + ggtitle("Sector Avg Prices") + theme(axis.text.x = element_text(angle = 90, hjust = 1))
```


Create summary data frame by Sector.
```{r}
sectorAvg = melt(finviz, id="Sector")
sectorAvg = subset(sectorAvg, variable %in% c("Price", "P.E", "PEG", "P.S", "P.B"))
sectorAvg = (na.omit(sectorAvg))
sectorAvg$value = as.numeric(sectorAvg$value)
sectorAvg = dcast(sectorAvg, Sector ~ variable, mean)
colnames(sectorAvg)[2:6] = c("SAvgPE", "SAvgPEG", "SAvgPS", "SAvgPB", "SAvgPrice")
```

Do the same summary process but by Industry instead of Sector.
```{r}
industryAvg = melt(finviz, id=c("Sector", "Industry"))
industryAvg = subset(industryAvg, variable %in% c("Price", "P.E", "PEG", "P.S", "P.B"))
industryAvg = (na.omit(industryAvg))
industryAvg$value = as.numeric(industryAvg$value)
industryAvg = dcast(industryAvg, Sector + Industry ~ variable, mean)
industryAvg = (na.omit(industryAvg))
colnames(industryAvg)[3:7] = c("IAvgPE", "IAvgPEG", "IAvgPS", "IAvgPB", "IAvgPrice")
```

Merge sector and industry average columns into original finviz dataset.
```{r}
finviz = merge(finviz, sectorAvg, by.x = "Sector", by.y = "Sector")
finviz = merge(finviz, industryAvg, by.x=c("Sector", "Industry"), by.y = c("Sector", "Industry"))
```

Create columns for determining over/underpricing based on comparison with other stocks in Sector and INdustry.
```{r}
finviz$SPEUnder <- 0
finviz$SPEGUnder <- 0
finviz$SPSUnder <- 0
finviz$SPBUnder <- 0
finviz$SPriceUnder <- 0
finviz$IPEUnder <- 0
finviz$IPEGUnder <- 0
finviz$IPSUnder <- 0
finviz$IPBUnder <- 0
finviz$IPriceUnder <- 0
```

Replace 0s with 1s where the value for the stock is less than the average to indicate that the stock may be undervalued based upon that metric.
```{r}
finviz$SPEUnder[finviz$P.E<finviz$SAvgPE] <- 1
finviz$SPEGUnder[finviz$PEG<finviz$SAvgPEG] <- 1
finviz$SPSUnder[finviz$P.S<finviz$SAvgPS] <- 1
finviz$SPBUnder[finviz$P.B<finviz$SAvgPB] <- 1
finviz$SPriceUnder[finviz$Price<finviz$SAvgPrice] <- 1
finviz$IPEUnder[finviz$P.E<finviz$IAvgPE] <- 1
finviz$IPEGUnder[finviz$PEG<finviz$IAvgPEG] <- 1
finviz$IPSUnder[finviz$P.S<finviz$IAvgPS] <- 1
finviz$IPBUnder[finviz$P.B<finviz$IAvgPB] <- 1
finviz$IPriceUnder[finviz$Price<finviz$IAvgPrice] <- 1
```

Sum the 10 columns to create a new column with index value from 1 to 10 indicating how undervalued stock is based on the different dimensions considered.

```{r}
finviz$RelValIndex = apply(finviz[79:88], 1, sum)
```

```{r}
underVal = subset(finviz, RelValIndex >= 8)
```

```{r}
head(underVal[, c("Ticker", "Company", "Country")])
```

Create a set of target stocks based on certain criteria.

```{r}
targetStocks = subset(finviz, Price > 20 & Price<100 & Volume>10000 & Country == "USA" 
            & EPS..ttm. > 0 & EPS.growth.next.year > 0 & EPS.growth.next.5.years>0 & Total.Debt.Equity < 1 & Institutional.Ownership < 30 & RelValInex > 8)
```

```{r}
targetStocks[, c("Ticker", "Company")]
```

